On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-22 Manuscript/manuscript.tex
Still Need to Address Some Comments

The goal of this paper is to provide recommendations on necessary sample size
for block bootstrap with dependent data, similar to what was done for basic
bootstrap in \citet{hesterberg2015teachers}. We consider a simple situation of
a stationary time series, where the parameters of interests are the mean,
standard deviation, and the first-order autocorrelation coefficient. We
compare six variants of block bootstrap CIs from the literature
\citep{diciccio1996bootstrap, rice2006mathematical}: a standard normal CI, and
a Student's $t$ CI, a percentile CI, a bias-corrected CI, a bias-corrected and
accelerated ($BC_a$) CI, and a CI proposed in this article. Their empirical
coverage rates at different sample sizes and dependence levels are compared in
a simulation study. The results of this study suggest that recovery of
temporal dependence parameters is reliant on the type of interval used.

@@

\paragraph{A Proposed CI}
\jy{Can we come up with a good name for the proposed method? Other people may cite this.}\mc{bias-centered? i'm not sure}
\jy{How about ``recentered percentile''? It has the same length as the
  percentile approach but is centered at the point estimate.}

We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$. A
$(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is asymmetric, as different critical values are used to compute
the lower and upper bounds.

@@

For estimating the mean parameter~$\mu$, Figure~\ref{fig:mu} suggests that all
methods eventually approach correct coverage of $\mu$ as sample size
increases. Student's $t$ CIs appear to need the smallest sample size
to achieve correct coverage, except for samples with strong negative
dependence, in which case, they actually over-cover $\mu$ for smaller sample
sizes. For instance, for a sample with $n = 100$ and $\phi = -0.4$, the lower
bound for a Students $t$ CIs coverage of $\mu$ is greater than 95\%, whereas
the coverage intervals for other methods contain 95\%. The standard,
Percentile, $BC$, and $BC_a$, and Proposed Method CIs require similar sample
sizes to recover $\mu$ at the nominal level for all combinations of $n$ and
$\phi$. All methods seem to require a smaller sample to recover $\mu$ at the
nominal rate when dealing with negative dependence versus positive dependence.
For example, $BC$ CIs recover $\mu$ for $n \geq 100$ when $\phi = -0.2$, but
they only recover $\mu$ for $n \geq 800$ when $\phi = 0.2$. In addition, as a
negative dependence gets stronger, holding everything else equal, coverage
increases, which lead to the Student $t$ CI's aforementioned over-coverage. As
a positive dependence gets stronger, holding everything else equal, coverage
decreases, and a larger sample is necessary to recover~$\mu$.

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that every method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods, and there is no over-coverage issue for
$\phi = -0.4$. Standard, percentile, $BC$, $BC_a$, and Proposed Method CIs
again have similar performance. All methods seem to have slightly higher
coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is positive.
Regardless of the sign, coverage of $\sigma_x$ gets worse as the strength of
the temporal dependence increases.
\jy{Compared to Figure 1, there is no over-coverage in this case. Say someting
  about this observation.}

For estimating the auto-correlation parameter~$\phi$, Figure~\ref{fig:phi}
suggests that while Standard, Student's $t$, and Proposed Method CIs do
approach correct coverage as sample size increases, Percentile, $BC$, and
$BC_a$ CIs deteriorate as sample size increases, especially as the strength of
the temporal dependence increases. Because of this, only Standard, Student's
$t$, and Proposed Method CIs should be considered as effective block bootstrap
methods to estimate $\phi$. Student's $t$ CIs once again can achieve correct
coverage with smaller sample sizes when compared to Standard and Proposed CIs
which perform similarly. Student's $t$ CIs can can recover $\phi$ at the
nominal level for $n \geq 100$ when the sample's temporal dependence is as
strong as 0.4. Coverage appears to be higher for all methods when the
dependence is negative rather than positive. Whether or not the dependence is
negative or positive, coverage of $\phi$ seems to increase slightly as the
absolute value increases.

\jy{Use cross for red. Otherwise, they are not distinguishable when printed in b/w.}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Proposed Method
    Bootstrap CIs for samples of $n = 800$ for an AR(1) process with
    $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs
    are displayed, as well as $\hat\theta_n$ (blue circle) and
    $\bar\theta_n^{(b)}$ (red cross). }
  \label{fig:npi}
\end{figure}

\jy{can you also do the same plot for sample size 1600? And then put the two
  plots in one figure?}

The outcomes of the $\phi$ estimation raise a natural question about the
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 100 CIs was generated for each of the percentile,
$BC$, and Proposed techniques. Illustrated in Figure~\ref{fig:npi}, it becomes
evident that the percentile-based CIs exhibit a notable bias, predominantly
manifesting as a substantial underestimation of $\phi$. On the other hand, $BC$
CIs do display a tendency to rectify this bias on average, albeit accompanied by
a rather inconsistent width across the intervals. Our proposed CIs manage to
effectively mitigate the bias, employing $\widehat{\text{bias}} =
\bar\theta_n^{(b)} - \hat\theta_n$ as a corrective measure. Notably, this
correction often takes a negative form, indicative of a positive adjustment
arising from the interval construction detailed in Section~\ref{sec:bbci}.

To summarize, the performance of the CIs depends on the target parameter. When
estimating $\mu$ and $\sigma_x$, any CI will do, although Student's $t$ CIs
perform noticeably better than the others. However, when estimating $\phi$,
the choice of method is of utmost importance as to avoid coverage
deterioration. A smaller sample size is generally required to estimate a
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the
same magnitude. In order to know if coverage will increase as the strength of
the temporal dependence increases, one need to know what the parameter of
interest is, and in the case of $\mu$, the direction of the serial dependence.
\jy{The BC approach appears not correcting the bias correctly. The proposed
  approach simply does not correct point estimate and only uses the spreadness
  from the bootstrap to construct CIs.}

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover
parameter of a time series at the nominal level given an infinitely large
sample, so the goal for this study was to find the smallest finite sample
length $n$ of a time series in order for the block bootstrap procedure to 
recover its associated parameters at an acceptable rate.
Our analysis relies on the assumption that there is a size $n$ large
enough for the method to work: that is, the method's performance improves as
$n$ increases. Out of the five types of intervals used in this study, this
assumption was found to hold true with respect to estimating $\phi$ only for
Standard, Student's $t$, and Proposed Method CIs, whereas Percentile, $BC$,
and $BC_a$ intervals exhibited coverage deterioration as $n$ increased. The
Percentile CI's coverage deterioration can be attributed to bias that is not
corrected as $n$ increases. Specifically, as $n$ increases, the width of the
CI decreases, but because the percentile CI underestimates $\phi$, the
coverage decreases. The $BC$ CI seems to correct the bias, but the width of
the CI seems to be inconsistent. The acceleration factor of the $BC_a$ CI
seems to fail, as the width of the CI seems to be inconsistent.

\jy{use a better topic sentence.}
When using Student's $t$ intervals and $\phi$ is unknown, the results of this
study suggest that an $n \geq 800$ may be necessary for common practice to
estimate simpler parameters like $\mu$ or $\sigma_x$. If $\phi$ is already
known, a lesser $n$ may be adequate to estimate these parameters. For lower
absolute values of $\phi$, a smaller $n$ is sufficient when estimating $\mu$
or $\sigma$. For negative values of $\phi$, coverage of $\mu$ is higher in
general. In fact, Student's $t$ seems to over-cover $\mu$ for $\phi = - 0.4$
when using smaller sample sizes, meaning that if $\phi$ is known to have a
very negative value, another CI may be required. However, in real world
applications, $\phi$ is more commonly found to be positive, so Student's $t$
should still be used if $\phi$ is unknown. Lastly, to estimate $\phi$,
$n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.

>>>>

The goal of this paper is to provide recommendations on necessary sample size
for block bootstrap with dependent data, similar to what was done for basic
bootstrap in \citet{hesterberg2015teachers}. We consider a simple situation of
a stationary time series, where the parameters of interests are the mean,
standard deviation, and the first-order autocorrelation coefficient. We
compare six variants of block bootstrap CIs from the literature
\citep{diciccio1996bootstrap, rice2006mathematical}: a standard normal CI, and
a Student's $t$ CI, a percentile CI, a bias-corrected CI, a bias-corrected and
accelerated ($BC_a$) CI, and a recentered percentile CI proposed in this article.
Their empirical
coverage rates at different sample sizes and dependence levels are compared in
a simulation study. The results of this study suggest that recovery of
temporal dependence parameters is reliant on the type of interval used.

@@

\paragraph{Recentered Percentile CI}
\jy{Can we come up with a good name for the proposed method? Other people may cite this.} \mc{bias-centered? i'm not sure}
\jy{How about ``recentered percentile''? It has the same length as the
  percentile approach but is centered at the point estimate.} \mc{replaced "proposed
  method" with "recentered"}

We propose a CI that is centered at the point estimator and uses the variation
in the bootstrap estimates to construct the error bound. This interval requires
the computation of $\bar\theta_n^{(b)}$, the mean of all bootstrap point
estimates, and $\widehat{\text{bias}} = \bar\theta_n^{(b)} - \hat\theta_n$. A
$(1 - \alpha)100\%$ CI is
\[
(\hat\theta_{n, \alpha/2}^{(b)} - \widehat{\text{bias}}, \quad
\hat\theta_{n, 1 - \alpha/2}^{(b)} - \widehat{\text{bias}}).
\]
This CI is centered around $\hat\theta_n$ and can also be rewritten
as 
\[
(\hat\theta_n + \hat\theta_{n, \alpha/2}^{(b)} - \bar\theta_n^{(b)}, \quad
\hat\theta_n + \hat\theta_{n, 1 - \alpha/2}^{(b)} - \bar\theta_n^{(b)}).
\]
It is not necessarily symmetric, as different critical values are used to compute
the lower and upper bounds. It has the same width as the Percentile CI.

@@

For estimating the mean parameter~$\mu$, Figure~\ref{fig:mu} suggests that all
methods eventually approach correct coverage of $\mu$ as sample size
increases. Student's $t$ CIs appear to need the smallest sample size
to achieve correct coverage, except for samples with strong negative
dependence, in which case, they actually over-cover $\mu$ for smaller sample
sizes. For instance, for a sample with $n = 100$ and $\phi = -0.4$, the lower
bound for a Students $t$ CIs coverage of $\mu$ is greater than 95\%, whereas
the coverage intervals for other methods contain 95\%. The Standard,
Percentile, $BC$, and $BC_a$, and Recentered CIs require similar sample
sizes to recover $\mu$ at the nominal level for all combinations of $n$ and
$\phi$. All methods seem to require a smaller sample to recover $\mu$ at the
nominal rate when dealing with negative dependence versus positive dependence.
For example, $BC$ CIs recover $\mu$ for $n \geq 100$ when $\phi = -0.2$, but
they only recover $\mu$ for $n \geq 800$ when $\phi = 0.2$. In addition, as a
negative dependence gets stronger, holding everything else equal, coverage
increases, which lead to the Student $t$ CI's aforementioned over-coverage. As
a positive dependence gets stronger, holding everything else equal, coverage
decreases, and a larger sample is necessary to recover~$\mu$.

For estimating the standard deviation parameter~$\sigma_x$,
Figure~\ref{fig:sigma} suggests that every method can reach nominal coverage of
$\sigma_x$ if the sample is large enough, but for a given $n$ and $\phi$,
coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general. Like
$\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample sizes
when compared to other methods. Unlike $\mu$, there is no over-coverage issue for
$\sigma$ when $\phi = -0.4$. Standard, Percentile, $BC$, $BC_a$, and Recentered
CIs again have similar performance. All methods seem to have slightly higher
coverage of $\sigma_x$ when $\phi$ is negative versus when $\phi$ is positive.
Regardless of the sign, coverage of $\sigma_x$ gets worse as the strength of
the temporal dependence increases.
\jy{Compared to Figure 1, there is no over-coverage in this case. Say someting
  about this observation.} \mc{addressed}

For estimating the auto-correlation parameter~$\phi$, Figure~\ref{fig:phi}
suggests that while Standard, Student's $t$, and Recentered CIs do
approach correct coverage as sample size increases, Percentile, $BC$, and
$BC_a$ CIs deteriorate as sample size increases, especially as the strength of
the temporal dependence increases. Because of this, only Standard, Student's
$t$, and Recentered CIs should be considered as effective block bootstrap
methods to estimate $\phi$. Student's $t$ CIs once again can achieve correct
coverage with smaller sample sizes when compared to Standard and Recentered CIs
which perform similarly. Student's $t$ CIs can can recover $\phi$ at the
nominal level for $n \geq 100$ when the sample's temporal dependence is as
strong as 0.4. Coverage appears to be higher for all methods when the
dependence is negative rather than positive. Whether or not the dependence is
negative or positive, coverage of $\phi$ seems to increase slightly as the
absolute value increases for Standard, Student's $t$, and Recentereed CIs. For the
values of $\phi$ observed, there are no examples
of over-coverage for Standard, Student's $t$, $BC$, $BC_a$, or Recentered CIs.
However, Percentile CIs appear to over-cover $\phi$ for smaller sample sizes when 
$\phi = -0.2$ and when $\phi = 0$, indicating again that they should not be used.

\jy{Use cross for red. Otherwise, they are not distinguishable when printed in b/w.}
\mc{addressed}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Recentered
    Bootstrap CIs for samples of $n = 800$ for an AR(1) process with
    $\phi = 0.4$. For each replicate, the lower and upper bounds of the CIs
    are displayed, as well as $\hat\theta_n$ (blue circle) and
    $\bar\theta_n^{(b)}$ (red cross). }
  \label{fig:npi}
\end{figure}

\jy{can you also do the same plot for sample size 1600? And then put the two
  plots in one figure?} \mc{addressed}

The outcomes of the $\phi$ estimation raise a natural question about the
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 100 CIs was generated for each of the percentile,
$BC$, and Recentered techniques. Illustrated in Figure~\ref{fig:npi}, it becomes
evident that the percentile-based CIs exhibit a notable bias, predominantly
manifesting as a substantial underestimation of $\phi$. On the other hand, $BC$
CIs do display a tendency to rectify this bias on average, albeit accompanied by
a rather inconsistent width across the intervals. Our Recentered CIs manage to
effectively mitigate the bias, employing $\widehat{\text{bias}} =
\bar\theta_n^{(b)} - \hat\theta_n$ as a corrective measure. Notably, this
correction often takes a negative form, indicative of a positive adjustment
arising from the interval construction detailed in Section~\ref{sec:bbci}.

To summarize, the performance of the CIs depends on the target parameter. When
estimating $\mu$ and $\sigma_x$, any CI will do, although Student's $t$ CIs
perform noticeably better than the others. However, when estimating $\phi$,
the choice of method is of utmost importance as to avoid coverage
deterioration. A smaller sample size is generally required to estimate a
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the
same magnitude. In order to know if coverage will increase as the strength of
the temporal dependence increases, one need to know what the parameter of
interest is, and in the case of $\mu$, the direction of the serial dependence. The
$BC$ approach does not seem to be correcting bias appropriately when estimating
$\phi$. Like the Percentile method, the Recentered method uses the spread from the 
bootstrap to construct the width of the CI. However, the Recentered approach, does
not correct from the original point estimate $\hat\theta_n$.
\jy{The BC approach appears not correcting the bias correctly. The Recentered
  approach simply does not correct point estimate and only uses the spreadness
  from the bootstrap to construct CIs.} \mc{addressed}

@@

Block bootstrap is a useful method for estimating parameters of a time series,
from simple parameters like the mean to more complicated temporal dependence
factors. We know theoretically that the block bootstrap procedure will cover
parameter of a time series at the nominal level given an infinitely large
sample, so the goal for this study was to find the smallest finite sample
length $n$ of a time series in order for the block bootstrap procedure to 
recover its associated parameters at an acceptable rate.
Our analysis relies on the assumption that there is a size $n$ large
enough for the method to work: that is, the method's performance improves as
$n$ increases. Out of the five types of intervals used in this study, this
assumption was found to hold true with respect to estimating $\phi$ only for
Standard, Student's $t$, and Recentered Method CIs, whereas Percentile, $BC$,
and $BC_a$ intervals exhibited coverage deterioration as $n$ increased. The
Percentile CI's coverage deterioration can be attributed to bias that is not
corrected as $n$ increases. Specifically, as $n$ increases, the width of the
CI decreases, but because the percentile CI underestimates $\phi$, the
coverage decreases. The $BC$ CI seems to correct the bias, but the width of
the CI seems to be inconsistent. The acceleration factor of the $BC_a$ CI
seems to fail, as the width of the CI seems to be inconsistent.

\jy{use a better topic sentence.}
One of the goals of this study was to provide practical recommendations for 
necessary sample sizes when using block bootstrap to estimate parameters of 
serially dependent data. When using Student's $t$ intervals and $\phi$ is unknown,
the results of this study suggest that an $n \geq 800$ may be necessary for common practice to
estimate simpler parameters like $\mu$ or $\sigma_x$. If $\phi$ is already
known, a lesser $n$ may be adequate to estimate these parameters. For lower
absolute values of $\phi$, a smaller $n$ is sufficient when estimating $\mu$
or $\sigma$. For negative values of $\phi$, coverage of $\mu$ is higher in
general. In fact, Student's $t$ seems to over-cover $\mu$ for $\phi = - 0.4$
when using smaller sample sizes, meaning that if $\phi$ is known to have a
very negative value, another CI may be required. However, in real world
applications, $\phi$ is more commonly found to be positive, so Student's $t$
should still be used if $\phi$ is unknown. Lastly, to estimate $\phi$,
$n \geq 100$ using the Student's $t$ method may be sufficient. Further
investigation may be necessary to see if there are other interval corrections
that fix the coverage deterioration problem for $\phi$.
